---
title: "R Notebook"
output: html_notebook
---
INFERENTIAL MODELLING – HDS 5320-02 
ASSIGNMENT – 06 (WEEK-06)
Vaishnavi Gannavaram
Instructor : Dr. Srikanth Mudigonda
---

```{r}
install.packages("MatchIt")
install.packages("optmatch")
```

```{r}
#Q1 Loading the dataset from package
require("faraway")
data("pima")
```

```{r}
#checking the summary
summary(pima)
```

```{r}
??pima
```

```{r}
View(pima)
```


```{r}
#Q2 . building full model using all predictors and outcome variable test to check relationship and then formulating the question. 
lmod.full <- glm(pima$test ~ pregnant  +glucose + diastolic+ triceps+ insulin + bmi+ diabetes+ age , family = binomial, data = pima)
summary(lmod.full)
```

#Based on each predictor's coefficients p values, predictors "pregnant","glucose","diastolic","bmi","diabetes" have significant relationship with the outcome variable.

Pima test.lmod.full = -8.40 + 0.12*pregnant + 0.03*glucose - 0.01*diastolic + 0.09*bmi +0.94*diabetes
---------------------------------------------------------
############ Formulating research question ###############

Does the presence of "test" variable affect the glucose levels, Considering the "diabetes", "BMI", "pregnant" as predictors.

Inorder to find out that , we are using "test" variable as treatment/controlled variable and glucose level as outcome variable with predictors bmi, diabetes, and pregnant. 
----------------------------------------------------------
```{r}
#Q3
#nearest method
library(MatchIt)
m.out1 <- matchit(test ~ bmi+diabetes+pregnant, 
                 data = pima,
                 method="nearest")
summary(m.out1)
plot(m.out1,type="jitter")
```
#From the above table, we can infer that
#Firstly, there are 268 observations which are controlled and treated among 500 observations. Hence, achieved balance.

#Secondly, there are 232 observations that did not have analogue on treated group. i.e, unmatched controlled observations. 

#Also, as there are fewer treated observations than controlled observations, the unmatched treated observations are 0.


#From the jitter plot, we can observe that
#The dispersion as the whole across the range of values is nearly identical in terms of how they fall in each range of of propensity scores.

```{r}
#optimal method
library(optmatch)
m.out.2 <- matchit(test ~ bmi+diabetes+pregnant, 
                 data = pima,
                 method="optimal")
summary(m.out.2)
plot(m.out.2, type="hist")
```
#Results from the optimal matching method were identical with the results from nearest matching method.
1. There are 268 observations which are controlled and treated among 500 observations. Hence, achieved balance.
2. There are 232 observations that did not have analogue on treated group. i.e, unmatched controlled observations. 
3. As there are fewer treated observations than controlled observations, the unmatched treated observations are 0.

#From the histogram plot, we can observe that the observations are evenly distributed.

Q4.
```{r}
#4a
#including treatment variable "test"
#install.packages("remotes")
#install.packages(
#"Zelig",repos = "http://r.iq.harvard.edu",type = "source")
library(remotes)
library(Zelig)
z.out <- zelig(glucose ~ bmi+diabetes+pregnant+test,
                 data = match.data(m.out1),
                 model = "ls")

x.out.0 <- setx(z.out, test=0)
x.out.1 <- setx(z.out, test=1)
s.out <- sim(z.out, x = x.out.0, x1 = x.out.1)
summary(s.out)
```
#comparing differences between controlled and treated groups with x being when treatment variable is 0 and x1 being treatment variable 1 including treatment variable "test".

#5
Two matched groups dont differ significantly in terms of the means of the outcomes between the treatment and control groups within a given matched dataset. 
95 % of confidence intervals overlap, we can conclude that they are not significantly different. 
Hence, presence of signs does not affect the glucose levels.

```{r}
#4b
#excluding treatment variable "test"
#install.packages("remotes")
#install.packages(
#"Zelig",repos = "http://r.iq.harvard.edu",type = "source")
library(remotes)
library(Zelig)
z.out <- zelig(glucose ~ bmi+diabetes+pregnant,
                 data = match.data(m.out1),
                 model = "ls")

x.out.0 <- setx(z.out)
x.out.1 <- setx(z.out)
s.out <- sim(z.out, x = x.out.0, x1 = x.out.1)
summary(s.out)
```
#5
Two matched groups dont differ significantly in terms of the means of the outcomes between the treatment and control groups within a given matched dataset. 
95 % of confidence intervals overlap, we can conclude that they are not significantly different. 




