---
title: "R Notebook"
output: html_notebook
---

---
INFERENTIAL MODELLING – HDS 5320-02 
ASSIGNMENT – 05 (WEEK-05)
Vaishnavi Gannavaram
Instructor : Dr. Srikanth Mudigonda
---


```{r}
#loading packages
library("glmnet")
library("faraway")
library("mice")
library("VIM")
library("lattice")
library("ggplot2")
library("dplyr")
library("tidyverse")
```


```{r}
#Loading the dataset from package
require("faraway")
data("dvisits")
```

```{r}
#checking the summary
summary(dvisits)
```

```{r}
help(dvisits)
```

```{r}
View(dvisits)
```

```{r}
#removing NA's
dvisits <- na.omit(dvisits)
```


#Q1
```{r}
#converting binary variables into as factors

dvisits$sex <- as.factor(dvisits$sex)
```

```{r}
dvisits$levyplus <- as.factor(dvisits$levyplus)
```

```{r}
dvisits$freepoor <- as.factor(dvisits$freepoor)
```

```{r}
dvisits$freerepa <- as.factor(dvisits$freerepa)
```

```{r}
dvisits$chcond1 <- as.factor(dvisits$chcond1)
```

```{r}
dvisits$chcond2 <- as.factor(dvisits$chcond2)
```

```{r}
#removed agesq to prevent multicollinearity
dvisitsedit <- select(dvisits, -agesq)
```

```{r}
summary(dvisitsedit)
```

#GRAPHS

```{r}
plot(dvisitsedit$doctorco, dvisitsedit$illness, pch = 19, col = "grey")

# Linear fit
abline(lm(dvisitsedit$doctorco ~ dvisitsedit$illness), col = "seagreen", lwd = 3)

```

The scatter plot with linear fit graph of outcome variable "doctorco" and predictor variable "illness" shows a positive and significant relationship between Number of consultations with a doctor or specialist in the past 2 weeks and Number of illnesses in past 2 weeks.

```{r}
plot(dvisitsedit$doctorco, dvisitsedit$age, pch = 19, col = "pink")

# Linear fit
abline(lm(dvisitsedit$doctorco ~ dvisitsedit$age), col = "blue", lwd = 3)

```
The scatter plot with linear fit graph of outcome variable "doctorco" and predictor variable "age" shows a positive and significant relationship between Number of consultations with a doctor or specialist in the past 2 weeks and age.


#Q2
Two predictors "chchond1" and "chcond2" are combined into a single predictor "chroniccond" with four levels corresponding to the combinations of both: 00, 01, 10, and 11.

```{r}
dvisitsedit$chroniccond <- as.factor(ifelse((dvisitsedit$chcond1 == 0 &
                                         dvisitsedit$chcond2 == 0),
                                        0,
                                 ifelse((dvisitsedit$chcond1 == 0 &
                                         dvisitsedit$chcond2 == 1),
                                        1,
                                 ifelse((dvisitsedit$chcond1 == 1 &
                                         dvisitsedit$chcond2 == 0),
                                        2, NA))))
```
#GRAPH

```{r}

plot(dvisitsedit$doctorco, dvisitsedit$chroniccond, pch = 19, col = "violet")

# Linear fit
abline(lm(dvisitsedit$doctorco ~ dvisitsedit$chroniccond), col = "green", lwd = 3)
```
The scatter plot with linear fit graph of outcome variable "doctorco" and predictor variable "chroniccond" shows a positive and significant relationship between Number of consultations with a doctor or specialist in the past 2 weeks and chronic conditions.

#Q3

```{r}
hist(dvisitsedit$doctorco)
#quick histogram of outcome variable "doctorco" will reveal its distribution
```
It is evident that a large proportion of the values are zeroes. Hence, fitting a "regular" Poisson count regression model on this dataset.

```{r}
fullmodel.poisson <- glm(doctorco ~ .- chcond1 - chcond2, family=poisson,
                   data=dvisitsedit)
summary(fullmodel.poisson)
```
"sex1", "freepoor1","illness","actdays","hospadmi","medicine","prescrib","old_female1" are the significant predictors found in this model with p value being less than 0.05. 

Here, we can see the null devience is 5634.8. and by adding a predictors to it the residual devience came out to be 4235.3. which signifies there is an increase in the chances to be be good fit. 


```{r}
print(paste("Deviance accounted for:",
            round((1-(fullmodel.poisson$deviance/fullmodel.poisson$null.deviance))*100, 3),
            "%", sep = ""))
```
We can observe that the difference between saturated and proposed models as 24.66%. The lower the devience the better the model. 24.66 being a lower devience we can say that the proposed model amounts to the good amount of variation for the data used. 

#Q4

```{r}
#install.packages("pscl")
library(pscl)
model.zip <- zeroinfl(doctorco ~ . - hospdays - nonpresc - chcond1 - chcond2, data=dvisitsedit)
summary(model.zip)
```
"prescrib","actdays","illness", "freerepa1","levyplus1" are noticed to be significant predictors in this model with p value being less than 0.05. 

"sex1","freepoor1","hospadmi","medicine","old_female1" are no longer significant as they were in the previous model.

#Q5
```{r}
better.model <- step(model.zip)
summary(better.model)
```
Among all the models, It is noticeable that the model with Lowest AIC i.e, AIC=6157.6
"doctorco ~ levyplus + freepoor + freerepa + illness + actdays + 
    hscore + hospadmi + medicine + prescrib + chroniccond " model can be the better fit.
 
#Q6
```{r}
dvisitsedit$young_male <- as.factor(ifelse((dvisitsedit$sex == 0 & 
                                     dvisitsedit$age <= 0.22),
                                  1, 0))
```

```{r}
dvisitsedit$old_female <- as.factor(ifelse((dvisitsedit$sex == 1 & 
                                   dvisitsedit$age >= 0.62),
                                1,0))
```


```{r}
mod.zip.reduced <- zeroinfl(doctorco ~ old_female, data=dvisitsedit2 )
summary(mod.zip.reduced)
```

```{r}
dvisitsedit2 <- select(dvisits, -agesq, -income,-freerepa,-chcond1,-chcond2,-nondocco,-nonpresc)

```


```{r}
x <- data.frame(old_female="1")
```

```{r}
predict(mod.zip.reduced, newdata=x,type="prob")
```
We can observe that the highest count probability is 0.7079 or 70.79%. lets calculate the probability from the zero-count part of our model.

```{r}
predict(mod.zip.reduced, newdata=x,type="zero")[1:8]
```
Firstly, NAs show up for all counts greater than 0, since our zero-count part of the model does not model counts greater than zero. The probability of producing zero is 0.4805297 or 48.05%. This is lesser than the value of 70.79% predicted by the Poisson-counts part of the model.

```{r}
mod.zip.reduced1 <- zeroinfl(doctorco ~ young_male, data=dvisitsedit2 )
summary(mod.zip.reduced1)
```
```{r}
y <- data.frame(young_male="1")
```

```{r}
predict(mod.zip.reduced1, newdata=y,type="prob")
```
We can observe that the highest count probability is 0.8831 or 88.31%. lets calculate the probability from the zero-count part of our model.


```{r}
predict(mod.zip.reduced1, newdata=y,type="zero")[1:8]
```

The probability of producing zero is 0.7960722 or 79.60%. This is lesser than the value of 88.31% predicted by the Poisson-counts part of the model.



```{r}
finalmodel.poisson <- glm(dvisitsedit$doctorco ~  dvisitsedit$prescrib + dvisitsedit$actdays + dvisitsedit$illness + dvisitsedit$old_female+ dvisitsedit$young_male , family=poisson,
                   data=dvisitsedit)
summary(finalmodel.poisson)
```


```{r}
ocount <- table(dvisitsedit$doctorco)[1:8]
pcount <- colSums(predprob(finalmodel.poisson)[,1:8])
ddf <- data.frame(ocount, round(pcount))
names(ddf) <- c("count","obs","pred")
plot(ddf$obs, ddf$pred, type="n", xlab="Observed counts", ylab="Predicted counts")
text(pcount,ocount, 0:7)
abline(0,1)
```
It is evident that a large proportion of the values are zeroes. Also, Observed counts and predicted counts seem to be fairly close (the numbers being close to the line with zero intercept and slope = 1).