---
title: "R Notebook"
output: html_notebook
---
```{r}
require("foreign")
require("ggplot2")
require("MASS")
require("Hmisc")
require("reshape2")
```


```{r}
library(survival)
data(pbc)
?pbc
summary(pbc)
View(pbc)
```

```{r}
my.pbc <- na.omit(pbc[, -1])
my.pbc$status <- as.factor(my.pbc$status) #status should be a factor
my.pbc$ascites <- as.factor(my.pbc$ascites) #ascites should be factor
my.pbc$stage <- as.factor(my.pbc$stage) #stage is ordinal/rank
summary(my.pbc)

```

```{r}
cor(my.pbc[, -c()])
```


```{r}
## ignoring trt, time, and status, creating a reduced dataset
my.reduced.pbc <- my.pbc[, -c(1:3)]
my.reduced.pbc$sex<- ifelse(my.reduced.pbc$sex=="m", 1, 0)
summary(my.reduced.pbc)
view(my.reduced.pbc)
```
## fit a model to predict "stage" using the other variables as predictors
```{r}
ormodel1= polr(stage ~ age  + sex + ascites + hepato + spiders +  edema +  bili + chol + albumin  + copper + alk.phos + ast + trig + platelet + protime , data = my.reduced.pbc, Hess = TRUE)
summary(ormodel1)
```

```{r}
# getting the p-values
# storing the coefficient table
ctable <- round(coef(summary(ormodel1)), 4)
# calculating and storing p-values
p <- pnorm(abs(ctable[, "t value"]), lower.tail = F) * 2
# combining coefficient table and p-values table
(ctable <- cbind(ctable, "p value" = round(p, 4)))
```

```{r}
# get confidence intervals
# log odd coefficients
or <- round(coef(ormodel1), 4)
# convert coefficients into odds ratio, combine with CIs
round(exp(cbind(OR = or)), 4)
```

```{r}
# get confidence intervals
# log odd coefficients
or <- round(coef(ormodel1), 4)
#ci
ci <- confint.default(ormodel1)
# convert coefficients into odds ratio, combine with CIs
round(exp(cbind(OR = or,ci)), 4)
```


```{r}
confint.default(ormodel1)
```

```{r}
library(brant)
brant(ormodel1)
```



```{r}
library(car)
vif(ormodel1)
```



```{r}
ormodel2= polr(stage ~bili + chol + copper + alk.phos + ast + trig, data = my.reduced.pbc, Hess = TRUE)
summary(ormodel2)
```

```{r}
# correlation plot
library(GGally)
xyz.var <- my.reduced.pbc[, c(7,8,10,11,12,13)]
ggpairs(xyz.var, title = "Correlation Plot between each Variable")
```

```{r}
library(brant)
brant(ormodel2)
```


```{r}
library(car)
vif(ormodel2)
```


```{r}
better.model <- step(ormodel2)
summary(better.model)
```


```{r}
vif(better.model)
```

```{r}
brant(better.model)
```






## identify relevant assumptions, and perform diagnostic tests to determine to what extent the assumptions were met.

```{r}
## Next, using my.pbc, create a new "event" variable to represent 0 and 1 as alive, and 2 as dead
## so 0 -> 0, 1 -> 0, 2 -> 1
my.pbc$newstatus <- as.numeric(ifelse(my.pbc$status == 2, 1, 0)) 
my.pbc <- my.pbc[, -2] ## remove the original status variable from the dataset
summary(my.pbc)
```
## fit a CoxPH model to predict the hazard function of newstatus

## state and test the assumptions of the CoxPH approach for the model fit above
## Is is reasonable to use CoxPH model, based on the diagnostic tests' results?

## Fit full-parametric models, where the survival function of status is modeled using
## exponential and Weibull distributions.
## Which of these is a better model?

## Comparing the two parametric models' outputs with that from the CoxPH model's
## output identify which predictors' relationship with the model's outcome
## remains the same, and which predictors' relationship changes, citing
## appropriate evidence.

## Taking all of the results together, which variables serve as reliable
## predictors of survival in the context of the dataset?
